On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-09-25 paper.tex
Writing stuff

\addbibresource{paper.bib}
\renewbibmacro*{volume+number+eid}{\ifentrytype{article}{\- \iffieldundef{volume}{}{Vol.~\printfield{volume},}\iffieldundef{number}{}{ No.~\printfield{number},}}}
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
\newbibmacro{string+doi}[1]{\iffieldundef{doi}{\iffieldundef{url}{#1}{\href{\thefield{url}}{#1}}}{\href{http://dx.doi.org/\thefield{doi}}{#1}}}
\DeclareFieldFormat[article, inproceedings, inbook, book, thesis]{title}{\usebibmacro{string+doi}{\mkbibquote{#1}}}
\renewcommand*{\bibfont}{\footnotesize}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\T}{T}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\colim}{colim}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\rk}{rk}
\newcommand{\catname}[1]{{\normalfont\textbf{#1}}}

@@

\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\XX}[1]{\mathcal{X}_{#1}}
\newcommand{\RR}[1]{\mathcal{R}_{#1}}

@@

Notice that for $\varpi \in P(\Phi^\vee)$ and $\beta \in \ZZ \Phi$ one has $(\varpi, \beta) \in \ZZ $.
Thus, for $\varepsilon \in R^*$ and $\varpi \in P(\Phi^\vee)$ we can define a character $\chi_{\varpi, \varepsilon} \in \Hom(\ZZ \Phi, R^*)$
 via the formula $\chi_{\varpi, \varepsilon}(\beta) = \varepsilon ^ {(\varpi, \beta)}$.

Every character $\chi \in \Hom(\ZZ \Phi, R^*)$ acts on the set of generators $\mathcal{X}_{\Phi, R}$ of the Steinberg group $\St(\Phi, R)$ via
\begin{equation} \chi \cdot x_\alpha(\xi) = x_\alpha(\chi(\alpha) \cdot \xi),\ \chi \in H,\ \alpha\in \Phi,\ \xi \in R. \end{equation}
It is not hard to check that the above formula gives a well-defined automorphism of $\St(\Phi, R)$ which we also denote by $\chi$.

\begin{example}
Let $A$ be a ring, denote by $R$ the ring $A[t, t^{-1}]$ of Laurent polynomials over $A$ and let $\alpha_i \in \Pi$ be a simple root.
Define automorphisms $\beta_i^+$ and $\beta_i^-$ of $\St(\Phi, R)$ via $\beta_i^+ = \chi_{\varpi_i, t}$, $\beta_i^- = \chi_{\varpi_i, t^{-1}}$.
It is easy to see that
\begin{equation}\label{eq:sigma_act} \beta_i^\pm(x_\alpha(\xi)) = x_\alpha(t^{\pm m_i(\alpha)} \cdot \xi).\end{equation}
\end{example}

\subsection{$\sigma$-pairs}

\begin{lemma} \label{lem:3.1f} Let $A$ be a local ring then any element $\alpha \in \St(\Phi, A)$ can be presented as a product 
$ $

\end{lemma}

@@

One of the key steps of in the proof of Suslin lemma for $K_2$ is to define analogue of maps $\beta_i^\pm$ on the level of the group $\St(\Phi, A[t])$.
Of course, one cannot expect such maps to be automorphisms or even be defined on the whole group $\St(\Phi, A[t])$.
Thus, our first goal is to define the subgroups of $\St(\Phi, A[t])$ on which these morphisms could, in principle, be defined.

@@

\begin{dfn} A pair of mutually inverse maps $(\sigma_i^+, \sigma_i^-)$ between the groups $N^i$ and $N_i$ will be called a {\it $\sigma$-pair corresponding to the vertex $i\in D(\Phi)$},
 if these maps fit into the following diagram.
\begin{equation} \label{eq:sigma-diagram}
\xymatrix{ N_i\ar@{-->}[r]^{\sigma_i^+}\ar[d] & N^i \ar[d] \ar@{-->}[r]^{\sigma_i^-} & N_i \ar[d] \\ 
          \St(\Phi, R) \ar@<-0.0ex>[r]_{\beta_i^+} & \St(\Phi, R) \ar@<-0.0ex>[r]_{\beta_i^-} & \St(\Phi, R)} 
\end{equation}
In the above diagram the vertical arrows are canonical embeddings of $N_k$ and $N^k$ into $\St(\Phi, A[t])$ postcomposed with 
 the map of Steinberg groups induced by the embedding $A[t] \hookrightarrow R = A[t, t^{-1}]$.
%As before, we denote by $R$ the ring $A[t, t^{-1}]$ of Laurent polynomials over $A$.
         % \GL(n, R) \ar@<-0.0ex>[r]_{(-)^{\delta_k}} & \GL(n, R) \ar@<-0.0ex>[r]_{{}^{\delta_k}(-)} & \GL(n, R)
\end{dfn}

>>>>

\addbibresource{paper.bib}
\renewbibmacro*{volume+number+eid}{\ifentrytype{article}{\- \iffieldundef{volume}{}{Vol.~\printfield{volume},}\iffieldundef{number}{}{ No.~\printfield{number},}}}
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
\newbibmacro{string+doi}[1]{\iffieldundef{doi}{\iffieldundef{url}{#1}{\href{\thefield{url}}{#1}}}{\href{http://dx.doi.org/\thefield{doi}}{#1}}}
\DeclareFieldFormat[article, inproceedings, inbook, book, thesis]{title}{\usebibmacro{string+doi}{\mkbibquote{#1}}}
\renewcommand*{\bibfont}{\footnotesize}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\T}{T}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\colim}{colim}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\rk}{rk}
\DeclareMathOperator{\Sym}{Sym}
\newcommand{\catname}[1]{{\normalfont\textbf{#1}}}

@@

\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\UU}{\hat{\mathrm{U}}}
\newcommand{\XX}[1]{\mathcal{X}_{#1}}
\newcommand{\RR}[1]{\mathcal{R}_{#1}}

@@

Notice that for $\varpi \in P(\Phi^\vee)$ and $\beta \in \ZZ \Phi$ one has $(\varpi, \beta) \in \ZZ $.
Thus, for $\varepsilon \in R^*$ and $\varpi \in P(\Phi^\vee)$ we can define a map of abelian groups $\chi(\varpi, \varepsilon) \in \Hom(\ZZ \Phi, R^*)$
 via the formula $\chi(\varpi, \varepsilon)(\beta) = \varepsilon ^ {(\varpi, \beta)}$.

Every map of abelian groups $\chi \in \Hom(\ZZ \Phi, R^*)$ specifies an action on the set of generators $\mathcal{X}_{\Phi, R}$ of the Steinberg group $\St(\Phi, R)$ via
\begin{equation} \chi \cdot x_\alpha(\xi) = x_\alpha(\chi(\alpha) \cdot \xi),\ \alpha\in \Phi,\ \xi \in R. \end{equation}
It is not hard to check that this action is compatible with Steinberg relations $\mathcal{R}_{\Phi, R}$ and 
 hence gives a well-defined automorphism of $\St(\Phi, R)$ which we denote by the same symbol $\chi$.

\begin{comment}
\begin{example}
Let $A$ be a ring, denote by $R$ the ring $A[t, t^{-1}]$ of Laurent polynomials over $A$.
Now if $\varpi \in P(\Phi^\vee)$ is a coweight then there is a pair of mutually inverse automorphisms $(\chi_{\varpi, t}, \chi_{-\varpi, t} = \chi_{\varpi, t^{-1}})$ 
 of $\St(\Phi, R)$ via $\beta_i^+ = \chi_{\varpi_i, t}$, $\beta_i^- = \chi_{\varpi_i, t^{-1}} = \chi_{-\varpi_i, t}$.
It is easy to see that
\begin{equation}\label{eq:sigma_act} \beta_i^\pm(x_\alpha(\xi)) = x_\alpha(t^{\pm m_i(\alpha)} \cdot \xi).\end{equation}
\end{example}

One of the key steps of in the proof of Horrocks theorem for $K_2$ is to define analogue of maps $\beta_i^\pm$ on the level of the group $\St(\Phi, A[t])$.
Of course, one cannot expect such maps to be automorphisms or even be defined on the whole group $\St(\Phi, A[t])$.
Thus, our first goal is to define the subgroups of $\St(\Phi, A[t])$ on which these morphisms could, in principle, be defined.
\end{comment}

Let $\varpi \in P(\Phi^\vee)$ be a co-weight of $\Phi$ as above.
Denote by $\mathcal{X}(\varpi)$ the subset of $\mathcal{X}_{\Phi, A[t]}$ consisting of those generators $x_{\alpha}(\xi)$ of $\St(\Phi, A[t])$ for which 
$(\alpha, \varpi) < 0$ implies that $\xi \in A[t]$ is divisible by $t^{-(\alpha, \varpi)}$.
% either $i\neq k$ (resp. $j\neq k$) or $\xi\in R[t]$ is divisible by $t$.
Denote by $N(\varpi)$ the subgroup of $\St(\Phi, A[t])$ generated by $\mathcal{X}(\varpi)$.

\begin{dfn}
Let $A$ be a ring and $R = A[t, t^{-1}]$ be the ring of Laurent polynomials over $A$.

By definition, a {\it $\sigma$-pair for $\varpi$} is a quadruple $(N(\varpi), N(-\varpi), \sigma(\varpi), \sigma(-\varpi))$ consisting of the following data:
\begin{itemize}
 \item a pair of subgroups $N(\varpi), N(-\varpi) \subseteq \St(\Phi, A[t])$;
 \item a pair of mutually inverse group-theoretic bijections $\sigma(\varpi)\colon N(\varpi) \to N(-\varpi)$, $\sigma(-\varpi)\colon N(-\varpi) \to N(\varpi)$
  fitting into the following commutative diagram:
\begin{equation} \label{eq:sigma-diagram}
\xymatrix{ N(\varpi) \ar@{-->}[r]^{\sigma(\varpi)}\ar@{^{(}->}[d] & N(-\varpi) \ar@{^{(}->}[d] \ar@{-->}[r]^{\sigma(-\varpi)} & N(\varpi) \ar@{^{(}->}[d] \\ 
          \St(\Phi, A[t]) \ar[d] & \St(\Phi, A[t]) \ar[d] & \St(\Phi, A[t]) \ar[d] \\
          \St(\Phi, R) \ar@<-0.0ex>[r]_{\chi(\varpi, t)} & \St(\Phi, R) \ar@<-0.0ex>[r]_{\chi(-\varpi, t)} & \St(\Phi, R).} \end{equation}  
\end{itemize}
\end{dfn}

\subsection{Factorization theorem} 
For a ring $R$ we denote by $\Sym(\Phi, R)$ the subgroup of $\St(\Phi, R)$ generated by Steinberg symbols $\{a, b\}_\alpha$, $a, b \in R^*$. 
 If $\rk(\Phi)\geq 2$ it is clear that $\Sym(\Phi, R)$ is a central subgroup of $\St(\Phi, R)$ contained in $\K_2(\Phi, R)$. 
Denote by $H(R) \subseteq \St(\Phi, R)$ the composite of the two subgroups $\Sym(\Phi, R)$ and $\UU(\Phi^+, R)$.
It is clear that $H$ is a functor on the category of commutative rings.

%Similarly $\Sym(\Phi, A[t, t^{-1}], M[t, t^{-1}]$ denotes the subgroup of $\St(\Phi, A[t, t^{-1}], M[t, t^{-1}])$ generated by elements of the form $h_\alpha(1+q)^{}$
As before $A$ is a commutative local ring with the maximal ideal $M$.

$j_+\colon A[t] \to A[t, t^{-1}]$. We write $j_+^*$ instead of $\St(\Phi, j_+)$.

Denote by $V_0$ the set $\St(\Phi, A[t]) \times H(A[t, t^{-1}]) \times \St(\Phi, A[t, t^{-1}], M[t, t^{-1}])$. 
We define an equivalence relation $\sim$ on $V_0$ as the minimal equivalence relation generated by the following 3 elementary transformations:
\begin{enumerate}
 \item $(a \cdot j_+^*(p), b, \beta) \sim (a, H(p) \cdot b, \beta)$, where $p$ is an element $H(A[t])$;
 \item $(a \gamma, b, \beta) \sim (a, b, ?(\gamma)^b \beta)$, where $\gamma$ is an element of $\St(\Phi, A[t], M[t])$;
\end{enumerate}

@@

@@

\begin{dfn} A pair of mutually inverse maps $(\sigma_i^+, \sigma_i^-)$ between the groups $N^i$ and $N_i$ will be called a {\it $\sigma$-pair corresponding to the vertex $i\in D(\Phi)$},
 if these maps fit into the following diagram.
In the above diagram the vertical arrows are canonical embeddings of $N_k$ and $N^k$ into $\St(\Phi, A[t])$ postcomposed with 
 the map of Steinberg groups induced by the embedding $A[t] \hookrightarrow R = A[t, t^{-1}]$.
%As before, we denote by $R$ the ring $A[t, t^{-1}]$ of Laurent polynomials over $A$.
         % \GL(n, R) \ar@<-0.0ex>[r]_{(-)^{\delta_k}} & \GL(n, R) \ar@<-0.0ex>[r]_{{}^{\delta_k}(-)} & \GL(n, R)
\end{dfn}
